Page not found

### 2023-07-28 manuscript/chapters/1_intro.qmd
fixed typos + improved eigenvalue plot

$$
\partial_t h(\vec z, t) = -h(\vec z, t) + \int_{\mathbb{R}^p} w(\vec z, \vec y) \phi(h(\vec y, t)) \rho(\mathrm d \vec y).
$$ {#eq-nf}
<!-- 
![(TODO add phi, add z instead of xi) The connectivity kernel defines the influence between localized neural populations in the embedding space](figures/connectivity_kernel.png) -->

@@

The main idea is to define a measurable bijective mapping $S : \Rp \mapsto [0,1]$ and a measure $\lambda : \Rp \mapsto \mathbb{R}^+$, such that we can write a one-dimensional neural field	$\tilde h(\alpha, t)$ with a connectivity kernel loosely defined as $\tilde w = w \circ S^{-1}$.

>>>>

$$
\partial_t h(\vec z, t) = -h(\vec z, t) + \int_{\mathbb{R}^p} w(\vec z, \vec y) \phi(h(\vec y, t)) \rho(\mathrm d \vec y).
$$ {#eq-nf}

In this equation, as well as for the rest of this work, we will use the notation $\rho(\mathrm d \vec y) := \rho(\vec y) \mathrm d \vec y$ to emphasize that the integral is weighted by the neuron density.

<!-- 
![(TODO add phi, add z instead of xi) The connectivity kernel defines the influence between localized neural populations in the embedding space](figures/connectivity_kernel.png) -->

@@

The main idea is to define a measurable bijective mapping $S : \Rp \mapsto [0,1]$ and a measure $\lambda : \Rp \mapsto \mathbb{R}^+$, such that we can write a one-dimensional neural field	$\tilde h(\alpha, t)$ with a connectivity kernel loosely defined as $\tilde w(\alpha, \beta) = w(S^{-1}(\alpha), S^{-1}(\beta))$.

### 2023-07-28 manuscript/chapters/2_neural_fields.qmd
added mean position, fixed typo
fixed typos + improved eigenvalue plot

The toy model we introduce is an instance of a low-rank neural network. Such models hold their name from the form of their connectivity matrix, which has a rank of (at most) $p$. We write the connectivity matrix as follows:

@@

The geometric view of these networks is that the recurrent currents lie in a $p$-dimensional subspace of $\R^N$ spanned by the vectors $\{\vec{F_1}, \cdots, \vec{F_\mu},\cdots, \vec{F_p}\}$ (where each $\vec{F_\mu}=(F_{\mu 1}, \cdots F_{\mu N})$), which therefore define a "natural" linear embedding (in the sense of @JazOst21) of the neural population activity.

@@

One last remark is that in our model, there is no external current; we only study the autonomous dynamics of the system.

@@

$$
\phi(h) = \frac{1}{1 + \mathrm{e}^h}.
$$

@@

We note that contrary to the model introduced in @schmutz2023convergence, we ignore the self-connections term, because it introduces a vanishing correction of order $\bO(1/N)$.

@@

```{python}
#| label: fig-spectrum
#| fig-cap: Spectrum of $K$ for $N=250$ and $p=2$. The numerical estimation (black, eigenvalues $\lambda_\mu$ and $\lambda_\perp=-1$) of the eigenvalues is close to the analytical ($N \to \infty$) derivation (red, only eigenvalues $\lambda_\mu$). Eigenvalues above the dotted gray line (positive values) correspond to unstable fixed points, and below (negative values) correspond to stable fixed points.
#| fig-subcap:
#|   - Pattern (stable) fixed point
#|   - Zero (unstable) fixed point
#| layout-ncol: 2
#| echo: false

dsigmoid = lambda x: sigmoid(x) * (1 - sigmoid(x))

def plot_spectrum(hstar_str: str):
	params = LowRankRNNParams.new_valentin(p=2, N=250, phi=sigmoid, I_ext=zero, exclude_self_connections=False)
	if hstar_str == 'pattern': hstar = params.F[:, 0]
	elif hstar_str == 'zero': hstar = np.zeros(params.N)
	dphi = dsigmoid(hstar)
	J = np.einsum('im,jm->ij', params.F, params.G * dphi[:, None]) / params.N - np.eye(params.N)
	eig_num = np.sort(np.linalg.eigvals(J).real)[::-1]
	eig_ana = { 'zero': 0.19061, 'pattern': -0.28090 }[hstar_str]  # see appendix A

	fig, ax = plt.subplots()
	mktrans(fig, ax)
	ax.axhline(0, color='gray', linestyle='--')
	ax.plot(eig_num[:8], marker='o', clip_on=False)
	ax.plot([7,8,8.5], eig_num[8:11], linestyle='--', color='k', clip_on=False)
	ax.plot(np.arange(params.p), np.full(params.p, eig_ana), color='r', marker='o', clip_on=False)
	ax.set_ylim((-1, 0.5))
	ax.set_xlim((0, 8.5))
	ax.set_ylabel('Eigenvalues')
	ax.set_xlabel('nth eigenvalue')
	breathe()
	return fig, ax

@@

$$
\begin{aligned}
\partial_t h(\vec z, t) &= -h(\vec z, t) + \int_{\Rp} w(\vec z, \vec y) \phi(h(\vec y, t)) \mathcal{N}^p(\d \vec y) \\
w(\vec z, \vec y) &= \sum_{\mu=1}^p z_\mu \tilde\phi(y_\mu) \\
\mathcal{N}^p(\d \vec y) &= \frac{\mathrm{exp}(\frac12\sum_{\mu=1}^p z_\mu^2)}{(2\pi)^{p/2}} \mathrm{d} y_1 \cdots \mathrm d y_p.
\end{aligned}
$$ {#eq-nf-gaussian-p}

@@

Due to the formulation of the neural field weighed by a probability distribution $\rho(y_1, \cdots, y_p)$, a natural way to estimate the integral is by Monte Carlo integration.
The essence of this method is to use the Central Limit Theorem, where we take $N$ independent samples $\vec{y_i} = (y_{1i}, \cdots, y_{pi})$ from the distribution $\rho$, and use them to estimate the integrand $I(\vec z, \vec y)$.

@@

We note however that @eq-nf-cycling-p no longer is a partial differential equation, but rather a delayed differential equation, which are in general much more complicated to solve. A simplifying assumption that we make is we extend the initial condition back in time, such that $h(\vec z, t < 0) = h(\vec z, 0)$ is a constant function. Numerically, this also means that we have to store a history of the neural fields instead of only the current one. Since the [`solve_ivp`](https://docs.scipy.org/doc/scipy/reference/generated/scipy.integrate.solve_ivp.html) method dynamically adapts the timestep of integration, we use linear interpolation between two known points in the stored neural field states to estimate $h(\vec z, t-\delta)$ (see [`lagging.py`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_lagging.py) and [its use](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_rnn.py#L334)).

>>>>

The toy model we introduce is an instance of a low-rank neural network. Such models hold their name from the form of their connectivity matrix, which has a rank $p$. We write the connectivity matrix as follows:

@@

The geometric view of these networks is that the recurrent currents lie in a $p$-dimensional subspace of $\R^N$ spanned by the linearly independant vectors $\{\vec{F_1}, \cdots, \vec{F_\mu},\cdots, \vec{F_p}\}$ (where each $\vec{F_\mu}=(F_{\mu 1}, \cdots F_{\mu N})$), which therefore define a "natural" linear embedding (in the sense of @JazOst21) of the neural population activity.

@@

In our model we make the simplifying assumption that there is no external current (we only study the autonomous dynamics of the system), but this does not hurt the generalizability of our results.

@@

$$
\phi(h) = \frac{1}{1 + \mathrm{e}^{-h}}.
$$

@@

We note that contrary to the model introduced in @schmutz2023convergence, which excludes the self-connection term by effectively setting $J_{ii} = 0$, we do not do this, because it introduces a vanishing correction of order $\bO(1/N)$.

@@

```{python}
#| label: fig-spectrum
#| fig-cap: Spectrum of $K$ for $N=1024$ and $p=5$. Numerical estimations (black, every line represents the sorted spectrum of one random sample of $K$, 20 samples total) of the eigenvalues are close to the analytical ($N \to \infty$) derivation (red, only eigenvalues $\lambda_\mu$). Eigenvalues above the dotted gray line (positive values) correspond to unstable fixed points, and below (negative values) correspond to stable fixed points. The first $p$ eigenvalues correspond to $\lambda_\mu$ ($\mu\in\{1,\cdots,p\}$), and the remaining eigenvalues correspond to $\lambda_\perp=-1$.
#| fig-subcap:
#|   - Pattern (stable) fixed point
#|   - Zero (unstable) fixed point
#| layout-ncol: 2
#| echo: false

N = 1024
p = 5
numeig = 20  # number of stored numerical eigenvalues
dsigmoid = lambda x: sigmoid(x) * (1 - sigmoid(x))

def plot_spectrum(hstar_str: str):
	nruns = 20
	eig_nums = np.zeros((nruns, numeig))

	for i, seed in enumerate(range(nruns)):
		params = LowRankRNNParams.new_valentin(p=p, N=N, phi=sigmoid, I_ext=zero, exclude_self_connections=False, random_state=seed)
		if hstar_str == 'pattern': hstar = params.F
		elif hstar_str == 'zero': hstar = np.zeros_like(params.F)
		dphi = dsigmoid(hstar)
		J = np.einsum('im,jm->ij', params.F, params.G * dphi) / params.N - np.eye(params.N)
		eig_nums[i] = np.sort(np.linalg.eigvals(J).real)[::-1][:numeig]

	eig_num = eig_nums.mean(axis=0)
	eig_num_std = eig_nums.std(axis=0, ddof=1) / np.sqrt(nruns)  # std of the mean estimator
	eig_ana = { 'zero': 0.19061, 'pattern': -0.28090 }[hstar_str]  # see appendix A

	fig, ax = plt.subplots()
	mktrans(fig, ax)
	# numerical estimations
	ax.axhline(0, color='gray', linestyle='--')
	# ax.plot(eig_num[:8], marker='o', clip_on=False)
	# ax.errorbar(np.arange(8), eig_num[:8], yerr=eig_num_std[:8], marker='o', clip_on=False, linestyle=':')
	ax.plot(eig_nums[:, :8].T, marker='.', color='k', alpha=0.2, clip_on=False, linestyle=':')
	ax.plot([7,8,8.5], eig_num[8:11], linestyle='--', color='k', clip_on=False)  # dashed line "fadeout"
	# analytical estimations
	ax.plot(np.arange(params.p), np.full(params.p, eig_ana), color='r', marker='o', linestyle=':', clip_on=False)
	ax.set_ylim((-1, 0.5))
	ax.set_xlim((0, 8.5))
	ax.set_ylabel('Eigenvalues')
	ax.set_xlabel('nth eigenvalue')
	breathe()
	return fig, ax

@@

$$
\begin{aligned}
\partial_t h(\vec z, t) &= -h(\vec z, t) + \int_{\Rp} w(\vec z, \vec y) \phi(h(\vec y, t)) \mathcal{N}^p(\d \vec y) \\
w(\vec z, \vec y) &= \sum_{\mu=1}^p z_\mu \tilde\phi(y_\mu) \\
\mathcal{N}^p(\d \vec y) &= \frac{\mathrm{exp}(\frac12\sum_{\mu=1}^p y_\mu^2)}{(2\pi)^{p/2}} \mathrm{d} y_1 \cdots \mathrm d y_p.
\end{aligned}
$$ {#eq-nf-gaussian-p}

@@

Due to the formulation of the neural field weighted by a probability distribution $\rho(y_1, \cdots, y_p)$, a natural way to estimate the integral is by Monte Carlo integration.
The essence of this method is to use the Central Limit Theorem, where we take $N$ independent samples $\vec{y_i} = (y_{1i}, \cdots, y_{pi})$ from the distribution $\rho$, and use them to estimate the integrand $I(\vec z, \vec y)$.

@@

We note however that @eq-nf-cycling-p no longer is a partial differential equation, but rather a delayed differential equation, which are in general much more complicated to solve. We make a simplifying assumption to extend the initial condition back in time, such that $h(\vec z, t < 0) = h(\vec z, 0)$ is a constant function. Numerically, the delayed partial differential equation also means that we have to store a history of the neural fields instead of only the current one. Since the [`solve_ivp`](https://docs.scipy.org/doc/scipy/reference/generated/scipy.integrate.solve_ivp.html) method dynamically adapts the timestep of integration, we use linear interpolation between two known points in the stored neural field states to estimate $h(\vec z, t-\delta)$ (see [`lagging.py`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_lagging.py) and [its use](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_rnn.py#L334)).
